---
title: "Number of true absent genes"
author: "Shankar K Shakya"
date: "October 20, 2016"
output: word_document
---


## Number of true absent genes
Previously I have produced heatmaps showing presence absence polymorphism. Here I am quantifying the number of absent genes, which is basically counting the number of NAs from presence absence polymorphism matrix. Beside that I have got rid of those genes that had no start and stop codon.

```{r, echo=TRUE, fig.align="center", fig.height=12, fig.width=12, global_options, echo=FALSE}

rm(list=ls())
library(vcfR)
library(ggplot2)

false_gene_mat_list <- list.files("FALSE GENE MATRICES/")
#false_gene_mat_list

pre_abs_mat_list <- list.files("PITG presence absence matrices/")
#pre_abs_mat_list

name_list <- strsplit(false_gene_mat_list, "_")
name_list <- sapply(name_list, function(x) x[1])
name_list

k <- 1
for (k in 1:length(pre_abs_mat_list)){
  
false_gene_mat <- read.table(file.path("FALSE GENE MATRICES/", false_gene_mat_list[k]))

pre_abs_mat <- as.matrix(read.table(file.path("PITG presence absence matrices/", pre_abs_mat_list[k])))

pre_abs_except_falsegenes <- setdiff(rownames(pre_abs_mat), false_gene_mat$LOCUS)

pre_abs_except_falsegenes_mat <- as.matrix(pre_abs_mat[rownames(pre_abs_mat) %in% pre_abs_except_falsegenes, ])


morder <- sort.int(apply(pre_abs_except_falsegenes_mat[,grep("mir|ipo|P7722", colnames(pre_abs_except_falsegenes_mat), invert = FALSE)], MARGIN=1, function(x){sum(is.na(x))}), index.return = TRUE)$ix

name <- name_list[k]

  heatmap.bp(pre_abs_except_falsegenes_mat[morder,], rlabels = FALSE)
  title(main = paste(name, "presence_absence_no_false_genes", nrow(pre_abs_except_falsegenes_mat),"/",nrow(pre_abs_mat)), outer = TRUE)

assign(paste(name, "presence_absence_no_false_genes", sep = "_"), as.matrix(pre_abs_except_falsegenes_mat))

}

cazy <- as.data.frame(apply(CAZy_presence_absence_no_false_genes, 2, function(x)
        length(which(is.na(x)))))
colnames(cazy) <- "cazy_P/A"

core <- as.data.frame(apply(Core_presence_absence_no_false_genes, 2, function(x) length(which(is.na(x)))))
colnames(core) <- "core_P/A"

crn <- as.data.frame(apply(CRN_presence_absence_no_false_genes, 2, function(x) length(which(is.na(x)))))
colnames(crn) <- "CRN_P/A"

effector <- as.data.frame(apply(Effector_presence_absence_no_false_genes, 2, function(x) length(which(is.na(x)))))
colnames(effector) <- "effector_P/A"

rxlr <- as.data.frame(apply(RxLR_presence_absence_no_false_genes, 2, function(x) length(which(is.na(x)))))
colnames(rxlr) <- "rxlr_P/A"

scr <- as.data.frame(apply(SCR_presence_absence_no_false_genes, 2, function(x) length(which(is.na(x)))))
colnames(scr) <- "scr_P/A"

elicitin <- as.data.frame(apply(elicitin_presence_absence_no_false_genes, 2, function(x) length(which(is.na(x)))))
colnames(elicitin) <- "elicitin_P/A"

pseudo <- as.data.frame(apply(pseudo_presence_absence_no_false_genes, 2, function(x) length(which(is.na(x)))))
colnames(pseudo) <- "pseudo_P/A"


npp <- as.data.frame(apply(NPP1_presence_absence_no_false_genes, 2, function(x) length(which(is.na(x)))))
colnames(npp) <- "NPP1_P/A"

prs_abs <- cbind(cazy, core, crn, effector, rxlr, scr, elicitin, pseudo, npp)
#prs_abs

absent_num <- apply(prs_abs, MARGIN = 1, sum)
head(absent_num)

split <- strsplit(names(absent_num), split = "_")
sample_names <- sapply(split, function(x) x[1])
sample_names[1:3] <- c("1306_1", "1306_2", "1306_3")
head(sample_names)

names(absent_num) <- sample_names
absent_num <- as.data.frame(absent_num)
absent_num <- cbind(sample_names, absent_num)
```

##testing only,
```{r}

head(absent_num)
absent_num_dist <- dist(as.matrix(absent_num), method = "manhattan")


plot(ape::nj(absent_num_dist))
```



## Plotting total number of absent genes per isolate and also by gene categories
```{r}

library("ggplot2")
library("forcats")

absent_plot <- ggplot(data = absent_num, aes(x = fct_inorder(sample_names), y = absent_num)) +
  geom_bar(stat = "identity") +
  theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = "Samples", y = "Number of absent genes") +
  theme(axis.text = element_text(size = 12)) 
  

absent_plot + coord_flip()


rownames(prs_abs) <- sample_names
prs_abs <- cbind(sample_names, prs_abs)
head(prs_abs) 


library(reshape)
prs_abs <- melt(prs_abs, id = c('sample_names'))

ggplot(prs_abs, aes(x=fct_inorder(sample_names), y = value)) + 
  geom_bar(stat = "identity") + facet_grid(variable~., scales = "free") +
   theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5)) +
  labs(x = "Samples", y = "Number of absent genes")


```


```{r, echo=FALSE, fig.align="center", fig.height=16, fig.width=16}

## Grouping of the samples 

head(absent_num)

mir <- absent_num[grep("mir|P7722", absent_num$sample_names), ]
ipo <- absent_num[grep("ipo", absent_num$sample_names), ]
pinf_mx <- absent_num[grep("inf|P10650", absent_num$sample_names), ]
pinf_clonal <- absent_num[grep("us|P10127|IN2009T1|RS2009P1|BL2009", absent_num$sample_names), ]
pinf_SA <- absent_num[grep("P13527|P13626", absent_num$sample_names), ]
pinf_Euro <- absent_num[grep("blue|P12204|NL07434", absent_num$sample_names),]
pinf_us1 <- absent_num[grep("P6096|LB|P11633|DDR", absent_num$sample_names),]


Pipo <- as.numeric(ipo$absent_num)
Pmir <- as.numeric(mir$absent_num)
Pinf_MX <- as.numeric(pinf_mx$absent_num)
Pinf_clonal <- as.numeric(pinf_clonal$absent_num)
Pinf_SA <- as.numeric(pinf_SA$absent_num)
Pinf_Euro <- as.numeric(pinf_Euro$absent_num)
Pinf_US1 <- as.numeric(pinf_us1$absent_num)

Pipo.df <- data.frame(Pipo, rep("P.ipo", length(Pipo)))
colnames(Pipo.df) <- c("absent", "Group")

Pmir.df <- data.frame(Pmir,rep("P.mir", length(Pmir)))
colnames(Pmir.df) <- c("absent", "Group")

Pinf_MX.df <- data.frame(Pinf_MX, rep("P.inf_MX", length(Pinf_MX)))
colnames(Pinf_MX.df) <- c("absent", "Group")

Pinf_clonal.df <- data.frame(Pinf_clonal,rep("P.inf_clonal", length(Pinf_clonal)))
colnames(Pinf_clonal.df) <- c("absent", "Group")

Pinf_SA.df <- data.frame(Pinf_SA, rep("P.inf_SA", length(Pinf_SA)))
colnames(Pinf_SA.df) <- c("absent", "Group")

Pinf_US1.df <- data.frame(Pinf_US1, rep("P.inf_US1", length(Pinf_US1)))
colnames(Pinf_US1.df) <- c("absent", "Group")

Pinf_Euro.df <- data.frame(Pinf_Euro, rep("P.inf_Euro", length(Pinf_Euro)))
colnames(Pinf_Euro.df) <- c("absent", "Group")

absent.df <- rbind(Pinf_US1.df, Pinf_clonal.df, Pinf_MX.df, Pinf_SA.df, Pinf_Euro.df, Pmir.df, Pipo.df)

ggplot(absent.df, aes(x=Group, y = absent)) + geom_boxplot(aes(fill = factor(Group))) +
  theme(axis.text.x = element_text(size = rel(2))) +
  theme(axis.text.y = element_text(size = rel(2))) + theme(strip.text = element_text(size = 25))

sig <- aov(absent ~ Group, absent.df)

summary(sig)


head(prs_abs)


mir <- prs_abs[grep("mir|P7722", prs_abs$sample_names),]
ipo <- prs_abs[grep("ipo", prs_abs$sample_names),]
pinf_mx <- prs_abs[grep("inf|P10650", prs_abs$sample_names),]
pinf_clonal <- prs_abs[grep("us|P10127|IN2009T1|RS2009P1|BL2009", prs_abs$sample_names),]
pinf_SA <- prs_abs[grep("P13527|P13626", prs_abs$sample_names),]
pinf_Euro <- prs_abs[grep("blue|P12204|NL07434", prs_abs$sample_names),]
pinf_us1 <- prs_abs[grep("P6096|LB|P11633|DDR", prs_abs$sample_names),]



Pmir.df <- data.frame(mir,rep("P.mir", length(mir)))
colnames(Pmir.df) <- c("samples", "Category", "Count", "Group")


Pipo.df <- data.frame(ipo,rep("P.ipo", length(ipo)))
colnames(Pipo.df) <- c("samples", "Category", "Count", "Group")


pinf_mx.df <- data.frame(pinf_mx,rep("P.inf_MX", length(pinf_mx)))
colnames(pinf_mx.df) <- c("samples", "Category", "Count", "Group")

pinf_clonal.df <- data.frame(pinf_clonal,rep("P.inf_clonal", length(pinf_clonal)))
colnames(pinf_clonal.df) <- c("samples", "Category", "Count", "Group")

pinf_SA.df <- data.frame(pinf_SA,rep("P.inf_SA", length(pinf_SA)))
colnames(pinf_SA.df) <- c("samples", "Category", "Count", "Group")

pinf_Euro.df <- data.frame(pinf_Euro, rep("P.inf_Euro", length(pinf_Euro)))
colnames(pinf_Euro.df) <- c("samples", "Category", "Count", "Group")

pinf_us1.df <- data.frame(pinf_us1, rep("P.inf_US1", length(pinf_us1)))
colnames(pinf_us1.df) <- c("samples", "Category", "Count", "Group")


all.df <- data.frame(rbind(pinf_us1.df, pinf_clonal.df, pinf_mx.df, pinf_SA.df, pinf_Euro.df, Pmir.df, Pipo.df))




# ggplot(all.df, aes(x=Group, y = Count)) + geom_point(aes(colour = factor(Group))) + facet_wrap(~Category, scales = "free")


library(ggplot2)
ggplot(all.df, aes(x=Group, y = Count)) + geom_boxplot(aes(fill = factor(Group))) + facet_wrap(~Category, scales = "free") + 
   theme(axis.text.x = element_text(size = rel(2), angle = 90, hjust = 1, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = rel(2))) +
  theme(strip.text = element_text(size = 25))

head(all.df)


ggplot(all.df, aes(x=Group, y = Count)) + facet_wrap(~Category, scales = "free") + geom_jitter(aes(colour = factor(Group), size = 3)) +
   theme(axis.text.x = element_text(size = rel(2), angle = 90, hjust = 1, vjust = 0.5)) +
  theme(axis.text.y = element_text(size = rel(2))) +
  theme(strip.text = element_text(size = 25))




```





















